Development of a 3D display

### 2016-05-18 paper.tex
gitignore, image creation

@@

@@

Something I had in mind since the beginning of this project were animated images. The Teensy 3.1
has \SI{256}{kB} of memory, which would fit 42 images of \SI{6}{kB} each. This
corresponts to roughly a second of playback. In order to display longer animations, there are
several workarounds: Instead of showing one image that spans the entirety of the rotation, I can
instead only save an image that is 25 pixels and display it four times.

@@

Another possibility would be video games. Although the limited resolution and computing power
would make it difficult to run the newest games, 3D versions of classics like Pong, Snake or a
simple platform game would be worth a shot.

>>>>

\usepackage{hyperref}

\usepackage[]{algorithm2e}

@@

\section{Image Creation}

Initially, all my images were made completely by hand. I had to enter the individual bits into the
program that would result in the desired image. Obviously this was a time-consuming and error
prone task. In order to simplify the process of making an image, I wrote a python program that
would support me in doing so. All of the source code is available on \url{github.com/mbjd/_3DPOV}.

\subsection{Basics}

In order to save an image, I used a three-dimensional array of 100 $\cdot$ 10 $\cdot$ 16 integers.
The lowest three bits of each integer represent the states of the red, green and blue LED
respectively. The rest of the bits are unused. This causes the program to use more memory than
strictly necessary, but it makes editing the pixels much simpler.

On the display itself everything stays the same, and no bits are wasted. The conversion from the
easily editable to the memory efficient format is automatically handled by the python program.

The pixels are arranged in a polar coordinate system. Because it's simpler to express straight
lines, spheres or surfaces in terms of cartesian coordinates, calculations typically go like this:

\begin{algorithm}[H]
\ForAll{pixels in the image}{
convert polar coordinates to cartesian\;
calculate some function with the coordinates as input\;
colour the pixel accordingly\;
}
\end{algorithm}

The actual source code of the program outlined below is fully available on
\url{http://github.com/mbjd/\_3DPOV} in the folder \texttt{image\_creation}.

\subsection{Straight lines}

Given two points, the program can draw a line connecting them. To do that, it calculates the
distance to the line for each point, and if it's below a certain threshold, the point is coloured.
This threshold controls the thickness of the line and can be selected freely, as can the colour of
the line. The implementation can be found in the functions \texttt{drawLine}.

\subsection{Spheres}

Spheres are even simpler than straight lines: For each point, the program compares the distance to
the center against the desired radius of the sphere. If it's smaller than the radius, the point is
within the sphere and can be coloured. This functionality is implemented as \texttt{drawSphere}.
Another version of this function, called \texttt{hollowSphere}, checks if the distance to the
middle point is within a certain range, drawing a hollow sphere.

\subsection{Planes}

A plane can be defined in terms of the equation $ax + by + cz = d$, where x, y, and z are the
cartesian coordinates, and a, b, c, and d are parameters defining the surface. Additionally, the
function takes an argument specifying the thickness of the plane. In the end, all pixels whose
coordinates $(x,y,z)$ satisfy $|ax + by + cz - d| < \frac{\si{thickness}}{2}$ are
coloured with the specified colour. The implementation is called \texttt{drawSurface}.

\subsection{Functions}

There are two fundamental types of functions that the program can draw. The first one are functions
of the form $f(x,y,z) \rightarrow colour$. In order to render such a function, I simply iterate
over all pixels in the image, calculating the function and colouring the pixel with its return
value.

Secondly, the program can plot functions of the form $f(x, y) \rightarrow z$. It is a special case
of the first one and could be defined in terms of it: I could simply iterate over all 16000 pixels
and check if $\lfloor f(x, y) \rfloor = z$. However, that would be inefficient, because the same
calculation would be made 10 times for the 10 different $z$ values. Instead, I only iterate over
all combinations of $x$ and $y$ values, and colour the pixel at the corresponding $z$ value. That
way, the function is only evaluated 1600 instead of 16000 times.

The respective implementations can be found in \texttt{plotColourFunction} and
\texttt{plotRealFunction}.

There's a third one called \texttt{plotBoolFunction}. It's similar to \texttt{plotColourFunction}
and can also be redefined in terms of it (with no significant loss in performance). Instead of
converting the coordinates $(x, y, z)$ to a colour, it takes a function that returns a boolean
value, and a colour. If the result of the given function is \texttt{True}, the pixel is coloured
with the given colour, otherwise nothing happens.

\subsection{Ada colourditional functionality}

Using these basic building blocks, I can draw more complex shapes by repeatedly calling one of the
above functions with different arguments. For example, there's the function
\texttt{connectcircle}. It takes a list of points and connects each with the next one, and the
last to the first point. That way I can draw arbitrary polygons.

I can also automatically draw a cuboid by specifying two points. The cuboid is then put up between
these points, all edges being parallel to the coordinate axes. The implementation is called
\texttt{drawCuboid}.

Another function, \texttt{connectAll}, takes a group of points and draws a line between each
possible pair of points.  That way, I can draw a 3-sided pyramid in one go.

@@

% Somehow \bit and \byte don't work in siunitx, so I'm just using the literal strings instead
Something I had in mind since the beginning of this project were animated images. The Teensy 3.1
has \SI{256}{kB} of memory, which would fit 42 images of \SI{6}{kB} each. This
corresponts to roughly a second of playback. In order to display longer animations, there are
several workarounds: Instead of showing one image that spans the entirety of the rotation, I can
instead only save an image that is 25 pixels and display it four times.

@@

Another fun thing to try would be video games. Although the limited resolution and computing power
would make it difficult to run the latest and greatest games, 3D versions of classics like Pong,
Snake or a simple platform game would certainly be worth a shot.
